---
title: "Extension of Villarreal Paper"
author: "Reisman and Lajous"
date: "April 23, 2015"
output: html_document
---

We chose to replicate Andrés Villarreal's 2004 AJS paper, "The Social Ecology of Rural Violence: Land Scarcity, the Organization of Agricultural Production, and the Presence of the State." Applying an ecological theory of crime to the study of rural violence in Mexico, Villarreal attempts explain the fact that rural municipalities have high variation in homicide rates by investigating the effects of land distribution, collective ownership, and the organization of agricultural production on homicide rates. He tests five specific hypotheses: (1) When agricultural land is scarce relative to the number of individuals, there will be more conflict and therefore more homicides; (2) An unequal distribution of land will lead to more violent conflict; (3) When property rights are not well enforced or are contingent, there will be more conflict over lands. For this reason, ejido and communal units will experience more violence; (4) agricultural production systems that involve more commodified relations of production and exchange will be associated with a breakdown of community social cohesion and therefore with more violence, and the introduction of cash crops will lead to greater conflict over resources and control over trade, and consequently to more violence; and (5) remote mountainous areas will have higher rates of violence due to the absence of state institutions.

We were unable to replicate Villarreal's findings, due in part to changes in Mexican data disclosure laws. For our extension of Villarreal's paper, we decided to update his analysis with newer (2007) data -- his original analysis was conducted using 1991 data. In doing so, we decided to remove the variables for which we identified data issues (The Thiel Index, Percent of Plots 5 Hectares or Less, and Log of Average Plot Size), as well as one variable for which 2007 data is not available (Percent Subsistence Agriculture). We correspondingly decided to include several new variables in the analysis, including a different indicator of land inequality to replace the Thiel index (the Gini Index), a new indicator of state presence not used by Villarreal (Number of Policemen per 1,000 people) and a new control variable -- the Human Development Index (HDI). We hope that these changes and additions will produce a valid analysis that we can compare to Villrreal's findings.


```{r, results = 'hide', warning=FALSE, error=FALSE}
library(ggplot2)
library(dplyr)
library(xlsx)
library(zoo)
library(foreign)
library(tidyr)
library(gdata)
library(maptools)
library(gpclib) 
library(RColorBrewer)
library(scales)
library(SDMTools)
library(readxl)
library(readr)
library(gdata)
```

**Adding New Variables**
```{r, warning=FALSE, error=FALSE}

#Poverty and Inequality (2010) 

##load CONEVAL poverty and inequality database
coneval <- read_csv("data/3.3 Concentrado, indicadores de pobreza por municipio.csv", skip = 6, col_names = FALSE)

##create table with muncode and percent in poverty, in extreme poverty and gini index 
poor <- coneval %>%
  filter(!is.na(X3)) %>%
  select(muncode = X3, pct.poor = X6, 
         pct.extpoor = X10, gini = X45) 

# Human development index (2010, per capita GNP, schooling, health) ------------
HDI <- read_csv("data/HDI_2010_UNDP.csv", skip = 3, 
                col_names = c("state", "mun", "Entidad",  "Municipio",  "school.level",
                              "expected.school",	"GNP.pc", 	"child.mort",	"education.index",
                              "income.index", 	"health.index",	"HDI"))
HDI <- HDI %>%
  mutate(muncode = state*1000 + mun) %>%
  select(muncode, HDI)

# police per 1000 pop -----------------------------------------------------
police <- read_csv("data/police_2010.csv", skip = 2, n_max = 2462, 
                   col_names = c("muncode", "state", "mun", "total.pers", "pol100k", "pol1k"))

police <- police %>%
  select(muncode, pol1k) %>%
  mutate(pol1k = ifelse(pol1k == "n. d.", NA, pol1k))
```

**Adding Updated Non-Agricultural Variables**
```{r, warning=FALSE, error=FALSE}

# doctors per 10000 pop ----------------------------------------------------
doctors <- read_csv("data/doctors_2005_2010.csv", skip = 2, n_max = 2456)

doctors <- doctors %>%
  select(muncode = Clave, docs = `2010`) %>%
  mutate(docs_10k = docs * 10) %>% 
  select(-docs) #making it docs per 10K to correspond with Villarreal

# homicides 2006, 2007, 2008 ----------------------------------------------
homicides <- read_csv(file="data/homicide_1990_2013_INEGI.csv", col_names = FALSE, skip = 6)

## new coloumn names, default were unreadable
colnames(homicides) <- c("muncode", "name", 2013:1990)
homicides[is.na(homicides)] <- 0

##clean out NAs and others, 
homicides <- homicides %>%
  filter(!grepl("996|997|998|991|993|992", muncode), muncode > 1000) %>%
  select(muncode, name, `2008`:`2006`) %>%
  mutate(hom_total = `2006` + `2007` + `2008`) %>%
  select(-`2006`, -`2007`, -`2008`, -name)

# census 2010 -------------------------------------------------------------
census <- read_delim("data/ITER_NALTXT10.TXT", delim = "\t", col_names = TRUE,
                     col_types = list(P3YM_HLI = col_numeric(), P8A14AN = col_numeric(), 
                                      P15YM_AN = col_numeric(), HOGJEF_F = col_numeric(),
                                      P_15A17_M = col_numeric(), P_18A24_M = col_numeric()))

census[is.na(census)] <- 0

## select relevant census variables
total_pop <- census %>%
  select(state = ENTIDAD, mun = MUN, mun.name = NOM_MUN, state.name = NOM_ENT, twn = LOC, pop = POBTOT,
         elev = ALTITUD, indi = P3YM_HLI, P8A14AN, P15YM_AN, fem.house = HOGJEF_F, total.house = TOTHOG,
         P_15A17_M, P_18A24_M) %>%
  mutate(muncode = state*1000 + mun,
         illiteracy = P8A14AN + P15YM_AN, 
         young.males = P_15A17_M + P_18A24_M) %>% #different categories in 2010 data
  select(-state, -mun, -P8A14AN, -P15YM_AN, -P_15A17_M, -P_18A24_M) %>%
  filter(twn != "0", twn != "9999", twn != "9998")

total_pop$total.house[total_pop$total.house == "*"] <- NA 
total_pop$total.house <- as.numeric(as.character(total_pop$total.house))

#population less than 2500 people
less_2500 <- census %>%
  filter(LOC != 0 & LOC != 9998 & LOC != 9999, POBTOT < 2500) %>%
  mutate(muncode = ENTIDAD*1000 + MUN) %>%
  group_by(muncode) %>%
  summarise(pop.less.2500 = sum(POBTOT))

total_pop <- total_pop %>%
  filter(!is.na(total.house)) #I think this shrinks the sample size a good amount - the data disclosure laws seem to have affected this too

## why do we need the total.house?

## all names have to be fixed so that they can be joined by name to the agr variables which don't have muncode.
total_pop$mun.name <- gsub("\xed", "i", total_pop$mun.name)
total_pop$mun.name <- gsub("\xfa", "u", total_pop$mun.name)
total_pop$mun.name <- gsub("\xf3", "o", total_pop$mun.name)
total_pop$mun.name <- gsub("\xe9", "e", total_pop$mun.name)
total_pop$mun.name <- gsub("\xe1", "a", total_pop$mun.name)
total_pop$mun.name <- gsub("\xfc\xbe\x98\xb6\x98\xbc", "u", total_pop$mun.name)
total_pop$mun.name <- gsub("\xc1", "A", total_pop$mun.name)
total_pop$mun.name <- gsub("\xfc\xbe\x8c\x96\x98\xbc", "n", total_pop$mun.name)

total_pop$state.name <- gsub("\xed", "i", total_pop$state.name)
total_pop$state.name <- gsub("\xfa", "u", total_pop$state.name)
total_pop$state.name <- gsub("\xf3", "o", total_pop$state.name)
total_pop$state.name <- gsub("\xe9", "e", total_pop$state.name)
total_pop$state.name <- gsub("\xe1", "a", total_pop$state.name)

total_pop$mun.name <- toupper(total_pop$mun.name)
total_pop$state.name <- toupper(total_pop$state.name)

total_pop$state.name <- gsub("COAHUILA DE ZARAGOZA", "COAHUILA", total_pop$state.name)
total_pop$state.name <- gsub("VERACRUZ DE IGNACIO DE LA LLAVE", "VERACRUZ", total_pop$state.name)
total_pop$state.name <- gsub("MICHOACAN DE OCAMPO", "MICHOACAN", total_pop$state.name)

total_pop <- total_pop %>%
  mutate(name = paste(mun.name, state.name, sep = ", "))

#separate out version to match naming errors
muns <- total_pop %>%
  group_by(muncode, name) %>%
  summarise(total.pop = sum(pop))

#subset out Oaxaca data
total.main <- filter(total_pop, !(muncode %in% 20001:20570))
total.oax <- filter(total_pop, muncode %in% 20001:20570)

#calculate weighed sd elevation
total.main <- total.main %>%   
  group_by(muncode, mun.name, state.name)  %>%
  summarise(total.pop = sum(pop), weigh.elev = wt.sd(elev, pop),  sd.elev = sd(elev),
            illiteracy = sum(illiteracy), indi = sum(indi), young.males = sum(young.males), 
            fem.house = sum(fem.house), total.house = sum(total.house))
            
#add all non-agricultural variables together into total_pop dataframe
total.main <- left_join(total.main, doctors, by = "muncode")
total.main <- left_join(total.main, HDI, by = "muncode")
total.main <- left_join(total.main, homicides, by = "muncode")
total.main <- left_join(total.main, police, by = "muncode")
total.main <- left_join(total.main, poor, by = "muncode")     
total.main <- left_join(total.main, less_2500, by = "muncode")
 #do we have a square kilometers area variable so we can calculate population density?  
```

**Adding Updated Agricultural Variables**
```{r, warning=FALSE, error=FALSE}
# agricultural variables --------------------------------------------------

## subsistance farming variable is not available

## total production units
prod_units <- read_excel("data/VIII Censo Agrícola 2007_Municipal.xls", sheet= "Cuadro 1", skip = 12,
                         col_names = FALSE) ## ignore crazy numbers in console

prod_units <- prod_units %>%
  filter(!is.na(X1), row_number() <= 2482, grepl("^\\s+|\\s+$", X1)) %>%
  mutate(state.name = ifelse(grepl("^\\s\\s+|\\s+$", X1), NA, X1))


prod_units$state.name <- na.locf(prod_units$state.name) ## subsistutes NAs with state names

prod_units <- prod_units %>%
  filter(grepl("^\\s\\s+|\\s+$", X1))

## property type, ejidos, comunal, privada

property_type <- read_excel("data/VIII Censo Agrícola 2007_Municipal.xls", 
                            sheet= "Cuadro 4", skip = 12, col_names = FALSE)


property_type  <- property_type %>%
  filter(!is.na(X1), row_number() <= 2482, grepl("^\\s+|\\s+$", X1)) %>%
  mutate(state.name = ifelse(grepl("^\\s\\s+|\\s+$", X1), NA, X1))


property_type$state.name <- na.locf(property_type$state.name) ## subsistutes NAs with state names

property_type <- property_type %>%
  filter(grepl("^\\s\\s+|\\s+$", X1))

# maize yields
corn <- read_excel("data/VIII Censo Agrícola 2007_Municipal.xls", 
                            sheet= "Cuadro 8", skip = 12, col_names = FALSE) 

corn <- corn %>%
  filter(row_number() <= 12730) %>%
  mutate(product = ifelse(!grepl("^\\s+|\\s+$", X1), X1, NA))

corn$product <- na.locf(corn$product)
corn$product <- gsub("`|\\'", "", iconv(corn$product, to="ASCII//TRANSLIT"))

corn <- corn %>%
  filter(grepl("^\\s+|\\s+$", X1)) %>% 
  mutate(state.name = ifelse(!grepl("^\\s\\s\\s+|\\s+$", X1), X1, NA))
corn$state.name <- na.locf(corn$state.name)

corn <- corn %>%
  filter(grepl("^\\s\\s\\s+|\\s+$", X1), grepl("MAIZ", product)) 

## units dedicated to coffee
coffee <-  read_excel("data/VIII Censo Agrícola 2007_Municipal.xls", 
                  sheet= "Cuadro 10", skip = 12, col_names = FALSE) 

coffee <- coffee %>%
  filter(row_number() <= 10448) %>%
  mutate(product = ifelse(!grepl("^\\s+|\\s+$", X1), X1, NA))

coffee$product <- na.locf(coffee$product)
coffee$product <- gsub("`|\\'", "", iconv(coffee$product, to="ASCII//TRANSLIT"))

coffee <- coffee %>%
  filter(grepl("^\\s+|\\s+$", X1)) %>% 
  mutate(state.name = ifelse(!grepl("^\\s\\s\\s+|\\s+$", X1), X1, NA))

coffee$state.name <- na.locf(coffee$state.name)


coffee <- coffee %>%
  filter(grepl("^\\s\\s\\s+|\\s+$", X1), grepl("CAFE", product)) 

## units dedicated to cattle
cattle <- read_excel("data/VIII Censo Agrícola 2007_Municipal.xls", 
                   sheet= "Cuadro 31", skip = 11, col_names = FALSE) 

cattle <- cattle %>%
  filter(row_number() <= 2481) %>%
  mutate(state.name = ifelse(!grepl("^\\s+|\\s+$", X1), X1, NA))

cattle$state.name <- na.locf(cattle$state.name)

cattle <- cattle %>%
  filter(grepl("^\\s+|\\s+$", X1))


#function to identify if the name is of a state or a municipality. If there are two leading white spaces it is a municipality, if there is only one white space it is a state and moved to a new column

naming <- function(df) {

df <- df %>%
  rename(mun.name = X1)

  ## standardizes state names
  df$state.name <- gsub("`|\\'", "", iconv(df$state.name, to="ASCII//TRANSLIT"))
  df$state.name <- gsub("COAHUILA DE ZARAGOZA", "COAHUILA", df$state.name)
  df$state.name <- gsub("VERACRUZ LLAVE", "VERACRUZ", df$state.name)
  df$state.name <- gsub("MICHOACAN DE OCAMPO", "MICHOACAN", df$state.name)
  
  ## removes accents and tilde in municipal names
  df$mun.name<- gsub("`|\\'", "", iconv(df$mun.name, to="ASCII//TRANSLIT"))
  df$mun.name <- gsub("~","", df$mun.name)
  df$mun.name <- gsub("¨","", df$mun.name)
  df$mun.name <- gsub("~","", df$mun.name)
  df$mun.name <- gsub("\"","", df$mun.name)
  
  
  df$mun.name <- trim(df$mun.name)
  df$state.name <- trim(df$state.name)
  df <- df %>%
    mutate(name = paste(mun.name, state.name, sep = ", "))
  
}

prod_units <- naming(prod_units)

prod_units <- prod_units %>%
  select(name, mun.name, state.name, total.units = X2, total.surface = X3)

property_type <- naming(property_type)

property_type <- property_type %>%
  select(name, mun.name, state.name, ejidal = X3, private = X5, comunal = X4)

corn <- naming(corn)

corn <- corn %>%
  select(name, mun.name, state.name, corn.yields = X6) %>%
  group_by(name, mun.name, state.name) %>%
  summarise(corn.yields = sum(corn.yields))

coffee <- naming(coffee)

coffee <- coffee %>%
  select(name, mun.name, state.name, coffee.units = X2)

cattle <- naming(cattle)

cattle <- cattle %>%
  select(name, mun.name, state.name, cattle.units = X2)

agr_var <- left_join(prod_units, property_type, by = c("name", "mun.name", "state.name"))
agr_var <- left_join(agr_var, corn, by = c("name", "mun.name", "state.name"))
agr_var <- left_join(agr_var, coffee, by = c("name", "mun.name", "state.name"))
agr_var <- left_join(agr_var, cattle, by = c("name", "mun.name", "state.name"))

#function correct_mun to change the of municipalities to match main data frame and agricultural variables 
#DO THIS FOR DATAFRAME OF AGRICULTURAL VARIABLES AFTER SUBSETTING OAXACA
errors <- read.csv("data/errors2007.csv", header = TRUE)

correct_mun <- function(x, pattern, replace) {
  for (i in seq_along(pattern))
    x <- gsub(pattern[i], replace[i], x, fixed = TRUE)
  x
}

#ag.var.main$name <- correct_mun(ag.var.main$name, errors$agricola2007, errors$censo2010)
#main.df <- left_join(total.main, ag.var.main, by = "name")
```

**Calculating variables for main data frame and filtering to main sample**
```{r, warning=FALSE, error=FALSE}
#once this is all done, calculate variables for main df
#main.df <- main.df %>%
#  mutate(prop.less.2500 = pop.less.2500/total.pop, pct.indi = (indi/total.pop) * 100, 
#        no.lit.rate = (illiteracy/total.pop) * 100, 
#         pct.fem.house = (fem.house/total.house) * 100,
#         pct.young = (young.males/total.pop) * 100,
#         log.pop.dens = log10(total.pop/sqkm),
#         hom.rate = hom_total/(total.pop*3) * 100000,
#         pct.ej = (ejidal/surface) * 100, 
#         pct.com = (comunal/surface) * 100, 
#         pct.individual = (private/surface) * 100,
#         log.corn.yield = log10(corn.tons/corn.ha), 
#         pct.cattle = (total.cattle.units/total.units) * 100,
#         pct.coffee = (total.coffee.units/total.units) * 100,
#         log.pop.surface = log10(total.pop/surface)) %>%
#  select(-twn, -illiteracy, -indi, -pop.less.2500, -fem.house, -total.house, -young.males, -sqkm, -hom_total, -ejidal, -surface, -comunal, -private, -total.units, -corn.tons, -corn.ha, -total.cattle.units, - total.coffee.units)

#Add dummy variable for State of Mexico, filter to main sample
#sample.main <- df.main %>%
#  mutate(dummy.SOM = as.numeric(muncode %in% 15001:15125))
#  filter(prop.less.2500 > .75)
```

**CONVERTING AND CALCULATING DATA FOR OAXACA**
```{r}
#subsetting elevation data because needs to be handled separately, then adding rest of non-agricultural variables to oaxaca dataset
oax.elev <- total.oax %>%
  select(name, muncode, elev, pop)
total.oax <- total.oax %>%
  select(-elev) %>%
  group_by(muncode, name) %>%
  summarise(total.pop = sum(pop),illiteracy = sum(illiteracy), indi = sum(indi), young.males = sum(young.males), 
            fem.house = sum(fem.house), total.house = sum(total.house))

total.oax <- left_join(total.oax, doctors, by = "muncode")
total.oax <- left_join(total.oax, HDI, by = "muncode")
total.oax <- left_join(total.oax, homicides, by = "muncode")
total.oax <- left_join(total.oax, police, by = "muncode")
total.oax <- left_join(total.oax, poor, by = "muncode")     
total.oax <- left_join(total.oax, less_2500, by = "muncode")

#correct errors in mun name for oaxaca agricultural variables, join this with total.oax
oax.errors <- read.csv("data/errors2007oax.csv", header = TRUE, stringsAsFactors = FALSE)
#fix name for a town that didn't match
oax.errors$censo2010[4] <- "SAN JUAN NUMI, OAXACA" #DO THIS ALSO WITH OAXACA AGR VAR DF

#combine census and agricultural dataframes
#oax.ag.var$name <- correct_mun(oax.ag.var$name, errors$agricola2007, errors$censo2010)
#oax.df <- left_join(total.oax, oax.ag.var, by = "name")

#oaxaca distritos 2010
oaxaca.distritos <- read.xlsx("data/oaxaca_distritos_2010.xls", 3, startRow = 6, endRow = 685, encoding = "latin1")
oaxaca.distritos <- tbl_df(oaxaca.distritos) 

#create table with municipality codes 
distritos <- oaxaca.distritos  %>%
  select(mun = Clave) %>%
  filter(!is.na(mun)) %>%
  mutate(mun = as.numeric(as.character(mun)), 
         muncode = (mun + 20000))

#create column that assigns districts to each muncode
distritos$distrito = rep(NA, nrow(distritos))
distritos$distrito[is.na(distritos$mun)] <- c(1:30)
distritos$distrito <- na.locf(distritos$distrito)

#filter out NA rows with district names
distritos <- distritos %>%
  filter(!is.na(mun))
```

**Calculating variables for Oaxaca**
```{r, warning=FALSE, echo=FALSE}
#join district table with oaxaca population table by district and generate new muncodes with district number. filter out new municipalies(districts) with more than 75% of population living in towns of less than 2500. 
oax.df$pop.less.2500[is.na(oax.df$pop.less.2500)] <- 0

oaxaca.d <- left_join(distritos, oax.df, by = "muncode") 
oaxaca.d <-filter(oaxaca.d, !(is.na(name))

#calculate variables for Oaxaca
oaxaca.d <- oaxaca.d  %>% 
  group_by(distrito) %>% 
  summarise(pop.less.2500 = sum(pop.less.2500), total.pop = sum(total.pop), prop.less.2500 = pop.less.2500/total.pop,
            indi = sum(indi), pct.indi = (indi/total.pop) * 100, 
            illiteracy = sum(illiteracy), no.lit.rate = (illiteracy/total.pop) * 100,
            sqkm = sum(sqkm), log.pop.dens = log10(total.pop/sqkm), 
            docs_10k = sum(docs_10k), 
            fem.house = sum(fem.house), total.house = sum(total.house), pct.fem.house = (fem.house/total.house) * 100,
            young.total = sum(young.total), pct.young = (young.total/total.pop) * 100, 
            hom_total = sum(hom_total), hom.rate = hom_total/(total.pop*3)*100000,
            ejidal = sum(ejidal), surface = sum(surface), log.pop.surface = log10(total.pop/surface), 
            pct.ej = (ejidal/surface) * 100,
            comunal = sum(comunal), pct.com = (comunal/surface) * 100,
            private = sum(private), pct.individual = (private/surface) * 100,
            total.cattle.units = sum(total.cattle.units), total.units = sum(total.units), 
            pct.cattle = (total.cattle.units/total.units) * 100,
            corn.tons = sum(corn.tons), corn.ha = sum(corn.ha), log.corn.yield = log10(corn.tons/corn.ha),
            total.coffee.units = sum(total.coffee.units), pct.coffee = (total.coffee.units/total.units) * 100, 
            pol1k = sum(pol1k)) %>%
  select(-illiteracy, -indi, -pop.less.2500, -fem.house, -total.house, -young.total, -sqkm, -hom_total, -ejidal, -surface, -comunal, -private, -total.units, -corn.tons, -total.cattle.units, -corn.tons, -corn.ha) %>%
  mutate(muncode = distrito + 20000)

oaxaca.dist$pct.coffee[is.na(oaxaca.dist$pct.coffee)] <- 0
 #do we have a square kilometers area variable so we can calculate population density?  

```

**Adding Weighted Variables for Oaxaca**
```{r, warning=FALSE, error=FALSE}
## SD ELEVATION WEIGHTED BY POPULATION 

## join population and elevation data frame with distritos in oaxaca 
pop_elev.dis <- left_join(oax.elev, distritos, by = "muncode")

## calculate sd elevation weighted by population grouped by distrito, and a dummy variable for the state of mexico
weigh_elev.oax <- pop_elev.dis %>% 
  group_by(distrito) %>% 
  summarise(weigh.elev = wt.sd(elev, pop)) %>% # use SDMTools to calculate weighted sd
  mutate(muncode = 20000 + distrito, dummy.SOM = 0) %>%
  select(-distrito)

## join weighted elevation with oaxaca main data frame
oaxaca.dist <- left_join(oaxaca.d, weigh_elev.oax, by = "muncode")

#calculate weighted SD of elevation for map
elev.map.oax <- left_join(distritos, oax.elev, by = "muncode") %>%
  group_by(distrito) %>%
  mutate(weigh.elev = wt.sd(elev, pop)) %>% 
  select(muncode, weigh.elev)

elev.map.oax <- elev.map.oax[!duplicated(elev.map.oax), ]

#GINI INDEX
#calculate population weighted average Gini index for Oaxacan municipalities
pop_gini <- left_join(total_pop, poor, by = "muncode")
pop_gini.dis <- left_join(pop_gini, distritos, by = "muncode")

weigh_gini.oax <- pop_gini.dis %>%
  group_by(distrito) %>%
  summarise(weigh.gini = wt.mean(gini, total.pop)) %>%
  mutate(muncode = 20000 + distrito) %>%
  select(-distrito)

weigh_gini.oax <- weigh_gini_oax %>%
  mutate(gini = weigh.gini) %>%
  select(-weigh.gini)

#join with main oaxaca dataframe
oaxaca.d <- left_join(oaxaca.d, weigh_gini.oax, by = "muncode")

#HDI
#calculate population weighted average HDI for Oaxacan municipalities
pop_hdi <- left_join(total_pop, HDI, by = "muncode")
pop_hdi.dis <- left_join(pop_hdi, distritos, by = "muncode")

weigh_hdi.oax <- pop_hdi.dis %>%
  group_by(distrito) %>%
  summarise(weigh.hdi = wt.mean(HDI, total.pop)) %>%
  mutate(muncode = 20000 + distrito) %>%
  select(-distrito)

weigh_hdi.oax <- weigh_hdi.oax %>%
  mutate(HDI = weigh.hdi) %>%
  select(-weigh.hdi)

#join with main oaxaca dataframe
oaxaca.d <- left_join(oaxaca.d, weigh_hdi.oax, by = "muncode")

#select variables for dataframe used in map for Oaxaca
oax.for.map <- select(oaxaca.d, muncode, hom.rate, sd.elev) 

#filter to oaxaca sample 
sample.oax <- oaxaca.d %>%  
  filter(prop.less.2500 > .75) 

```

**Bind two subsets into main sample**
```{r, warning=FALSE, error=FALSE}
sample.main <- sample.main %>%
  select(muncode, total.pop, dummy.SOM, prop.less.2500, pct.indi, no.lit.rate, docs_10k, pct.fem.house, pct.young, log.pop.dens, hom.rate, pct.ej, pct.com, pct.individual, log.corn.yield, pct.cattle, sd.elev, log.pop.surface,  pct.coffee, gini, pol1k, HDI)

sample.oax <- sample.oax %>%
  select(muncode, total.pop, dummy.SOM, prop.less.2500, pct.indi, no.lit.rate, docs_10k, pct.fem.house, pct.young, log.pop.dens, hom.rate, pct.ej, pct.com, pct.individual, log.corn.yield, pct.cattle, sd.elev, log.pop.surface,  pct.coffee, gini, pol1k, HDI)

sample <- rbind(sample.main, sample.oax)


#fix log corn yield, which was somehow coded as a character string
sample$log.corn.yield <- as.numeric(sample$log.corn.yield)
sample$log.pop.surface <- as.numeric(sample$log.pop.surface)
```

**Results**
```{r, results = 'hide', warning=FALSE, error=FALSE}
library(MASS)
library(texreg)
```

```{r, warning=FALSE, error=FALSE}
#Create function to extract necessary goodness-of-fit statistics for table, such that it matches Villarreal's table
extract.negbin <- function(model) {
  s <- summary(model)
  names <- rownames(s$coef)
  co <- s$coef[, 1]
  se <- s$coef[, 2]
  pval <- s$coef[, 4]
  
  th <- 1/(s$theta)
  se.th <- s$SE.theta
  ll <- (s$twologlik)/2
  n <- nobs(model)
 
  gof <- c(th, se.th, n, ll)
  gof.names <- c("Overdispersion parameter alpha", "Standard error of overdispersion", "N", "Log Likelihood")
 
  tr <- createTexreg(
  coef.names = names,
  coef = co,
  se = se,
  pvalues = pval,
  gof.names = gof.names,
  gof = gof
  )
  return(tr)
  
  }
```

*2 Models for Hypotheses 1 and 2:*
(1) When agricultural land is scarce relative to the number of individuals, there will be more conflict and therefore more homicides
(2) An unequal distribution of land will lead to more violent conflict


```{r, warning=FALSE, error=FALSE}
#Run models and build tables
One.a <- glm.nb(hom.rate ~ gini + log.corn.yield + log.pop.dens + pct.young + pct.indi + no.lit.rate + HDI + pct.fem.house + docs_10k + dummy.SOM, data = sample) 

One.b <- glm.nb(hom.rate ~ pop.per.HA + gini + log.corn.yield + log.pop.dens + pct.young + pct.indi + no.lit.rate + HDI + pct.fem.house + docs_10k + dummy.SOM,data = sample) 
```

```{r, results='asis'}
gsub("\n[[:space:]]","", htmlreg(list(extract.negbin(One.a), extract.negbin(One.b)), caption = "Table 1: Coefficients from the Negative Binomial Regression Models of Homicide on Land Scarcity and Inequality (Hypotheses 1 and 2)", caption.above = TRUE, stars = c(.05, .01), custom.coef.names = c("Constant", "Persons per hectare", "Gini index", "Log maize yields", "Log population density", "% young males", "% indigenous", "Illiteracy rate", "Human Development Index", "% female-headed households", "Medical doctors per 10,000", "Valley of Mexico"), custom.note = "SEs in parentheses,         * P < .05, ** P < .01, two-tailed tests", include.intercept = TRUE, reorder.coef = c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 1), groups = list("Plot size and crop yields" = 1:3, "Demographic variables" = 4:6, "Resource deprivation" = 7:8, "Family structure" = 9, "Access to medical attention" = 10, "Geographical region" = 11), inline.css = TRUE, doctype = FALSE, html.tag = FALSE, head.tag = FALSE, body.tag = FALSE, center = FALSE, digits = 4, star.symbol = "\\*")) 
```

*3 Models for Hypotheses 3 and 4:*
(3) When property rights are not well enforced or are contingent, there will be more conflict over lands. For this reason, ejido and communal units will experience more violence
(4) agricultural production systems that involve more commodified relations of production and exchange will be associated with a breakdown of community social cohesion and therefore with more violence, and the introduction of cash crops will lead to greater conflict over resources and control over trade, and consequently to more violence


```{r, warning=FALSE, error=FALSE}
Two.a <- glm.nb(hom.rate ~ pct.ej + pct.com + gini + log.corn.yield + log.pop.dens + pct.young + pct.indi + no.lit.rate + HDI + pct.fem.house + docs_10k + dummy.SOM, data = sample)
 
Two.b <- glm.nb(hom.rate ~ pct.ej + pct.com + pct.individual + pct.coffee + pct.cattle + gini + log.corn.yield + log.pop.dens + pct.young + pct.indi + no.lit.rate + HDI + pct.fem.house + docs_10k + dummy.SOM, data = sample) 
 
Two.c <- glm.nb(hom.rate ~ pct.ej + pct.com + pct.coffee + pct.cattle + gini + log.corn.yield + log.pop.dens + pct.young + pct.indi + no.lit.rate + HDI + pct.fem.house + docs_10k + dummy.SOM, data = sample) 
```

```{r, results='asis'}
gsub("\n[[:space:]]","", htmlreg(list(extract.negbin(Two.a), extract.negbin(Two.b), extract.negbin(Two.c)), caption = "Table 2: Coefficients from the Negative Binomial Regression Models of Homicide on Collective Ownership, Agricultural Organization and Production of Cash Crops (Hypotheses 3 and 4)", caption.above = TRUE, stars = c(.05, .01), custom.coef.names = c("Constant", "% ejido surface area", "% communal surface area", "% individual production", "% coffee production", "% cattle production", "Gini index", "Log maize yields", "Log population density", "% young males", "% indigenous", "Illiteracy rate", "Human Development Index", "% female-headed households", "Medical doctors per 10,000", "Valley of Mexico"), custom.note = "SEs in parentheses,         * P < .05, ** P < .01, two-tailed tests", include.intercept = TRUE, reorder.coef = c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 1), groups = list("Collective ownership" = 1:2, "Organization of agricultural production" = 3:5, "Plot size and crop yields" = 6:7, "Demographic variables" = 8:10, "Resource deprivation" = 11:12, "Family structure" = 13, "Access to medical attention" = 14, "Geographical region" = 15), inline.css = TRUE, doctype = TRUE, html.tag = FALSE, head.tag = FALSE, body.tag = FALSE, digits = 4, center = FALSE))
```

*1 Model for Hypothesis 5:* 
(5) remote mountainous areas will have higher rates of violence due to the absence of state institutions


```{r, warning=FALSE, error=FALSE}
Three <- glm.nb(hom.rate ~ weigh.elev + pol1k + pct.ej + pct.com + pct.individual + pct.coffee + pct.cattle + gini + log.corn.yield + log.pop.dens + pct.young + pct.indi + no.lit.rate + HDI + pct.fem.house + docs_10k + dummy.SOM, data = sample) 
```

```{r, results = 'asis'}
gsub("\n[[:space:]]","", htmlreg(extract.negbin(Three), caption = "Table 3: Coefficients from the Negative Binomial Regression Models of Homicide on State Presence (Hypothesis 5)", caption.above = TRUE, stars = c(.05, .01), custom.coef.names = c("Constant", "SD of altitude", "Police per 1,000" "% ejido surface area", "% communal surface area", "% individual production", "% coffee production", "% cattle production","Gini index", "Log maize yields", "Log population density", "% young males", "% indigenous", "Illiteracy rate", "Human Development Index", "% female-headed households", "Medical doctors per 10,000", "Valley of Mexico"), custom.note = "SEs in parentheses,         * P < .05, ** P < .01, two-tailed tests", include.intercept = TRUE, reorder.coef = c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 1), groups = list("State presence" = 1:2, "Collective ownership" = 3:4, "Organization of agricultural production" =5:7, "Plot size and crop yields" = 8:9, "Demographic variables" = 10:12, "Resource deprivation" = 13:14, "Family structure" = 15, "Access to medical attention" = 16, "Geographical region" = 17), inline.css = TRUE, doctype = FALSE, html.tag = FALSE, head.tag = FALSE, body.tag = FALSE, center = FALSE, digits = 4, star.symbol = "\\*")) 
```

**Conclusions**